---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# keylist

<!-- badges: start -->
<!-- badges: end -->

Lightweight `list` extensions that enforce unique keys.

## Installation

You can install the development version of keylist like so:

``` r
# install.packages("pak")
pak::pak("LJ-Jenkins/keylist")
```

## Overview

keylist provides two lightweight keylist S3 classes `klist` and `knlist`: extensions to `list` that enforce unique keys. `klist` accepts unnamed and named elements whilst `knlist` only accepts named elements.

```{r intro, error=TRUE}
library(keylist)

# klist - unnamed or named (if names present, they must be unique).
klist(1, a = 2)
klist(a = 1, a = 2)

# knlist - fully named and unique.
knlist(a = 1, b = 2)
knlist(1, b = 2)
knlist(a = 1, a = 2)

# keylist as more general function.
class(keylist(1))
class(keylist(a = 1, .named = TRUE))
```

As extensions to `list`, keylists can take any input list can. Inputs lists are not validated as they are not keylists - keylist validation operates on its own level so for validation of nested elements they must also be keylists.

```{r nest, error=TRUE}
# input list separate entity.
keylist(1, list(a = 1, a = 1))

# each keylist validates its own level.
keylist(1, keylist(a = 1, a = 1))
```

All nested elements of a list can be recursively turned into a keylist
when using one of the `as.*` variants.

```{r recurse, error=TRUE}
x <- list(1, list(a = 1, a = 1))
as.keylist(x, .recursive = TRUE)
```

## Assignment and Modification

Although lists in R can accept duplicate names, R preferentially chooses the existing list's name structure when it comes to assignment. Therefore if you can guarantee that the keys are unique at creation, there is no way of assigning a duplicate key using the normal R methods of `[`, `[[` and `$`.  See this example:

```{r list example, error=TRUE}
x <- list(1, a = 2, 3, b = 4)
x[1] <- list(x = 1)
x[1] # remains unnamed.

x["a"] <- list(x = 1)
x[2] # name remains 'a'.

x[5] <- list(k = 10) # index 5 doesn't exist...
x[5] # yet still remains unnamed.
```

keylist makes use of the above by not implementing any special validation tricks - it simply ensures the keys are unique upon creation then let's you go on your way.

However, assigning by position could enable a non-named element for a `knlist`. To circumvent this (and to give a clearer definition of type), `knlist` objects only allow assignment when the indexing is done by name.

```{r knlist assignment, error=TRUE}
x <- knlist(a = 1)
x[[1]] <- 1

# extraction by index is allowed.
x[[1]]
```

After creation, the main way that duplicate names may creep into lists is with direct name editing with `names()` or with list concatenation with `c()`. keylist includes S3 methods for both `klist` and `knlist` for these functions to ensure that duplicate keys are not added from these methods.

```{r names and c, error=TRUE}
x <- klist(1, a = 1)
names(x) <- c("a", "a")

x <- knlist(a = 1)
# name removal allowed for klist, but not knlist.
names(x) <- NULL

# knlist doesn't accept "" or NA names (like klist does).
setNames(x, NA)

class(c(klist(1, a = 2), 2, list(3)))
c(klist(1, a = 2), a = 2, list(3))
```

## Getting help

If you encounter a clear bug, please file an issue with a minimal reproducible example on [GitHub](https://github.com/LJ-Jenkins/keylist/issues).

## Code of Conduct

Please note that the keylist project is released with a [Contributor Code of Conduct](https://lj-jenkins.github.io/keylist/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
